gss marriage politics analysis

### 2023-04-20 outputs/paper/paper.qmd
fixing table titles

variable_table <-
  tibble(
    "Variable" = c(
      "partyid",
      "polviews",
      "marital",
      "sex",
      "cohort"
    ),
    "Variable Description" = c(
      "Self-ascribed belonging to the Republican, Democratic, or Indepedent parties",
    "Self-ascribed placement on the political spectrum from 'extremely liberal to extremely conservative'",
    "Marital status of respondent",
    "Self-ascribed sex of respondent",
    "Generational cohort that respondent belongs to "
    )
  )

@@

### Marital Status and Birth Cohort

Another important point of bias to acknowledge is the effect of birth cohort on marital status. As can be seen in @fig-1, the effects of age on birth cohorts largely impacts the marital status. For example, in 2014, there are significantly more respondent's belonging to the "Post-War" cohort and no respondent's belonging to the "Gen Z" cohort. This is influential as the 2014 data is missing marital rates from the "Gen Z" cohort as they were not old enough to be surveyed, thereby distorting marital rates in that year when compared to the follow three years. Additionally, due to the young age of "Gen Z", respondents many have not reached the age where they are looking for long-term relationships resulting in very few being married and even less being divorced and widowed, as shown in @fig-1. However, it is important to include them in this analysis as they are a growing and extremely relevant cohort in demographic studies. Additionally, it is noticeable that trends in marriage between birth cohorts are shifting as can be seen by the increase of "Never Married" respondents between cohorts each year. This is significant as it may act as a co-indicator of marital rates in the U.S.

```{r}
#| include: false
#| message: false
#### Creating Graph for comparison 2021 respondent's birth cohort against marital status ####
cohort_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false
#### Creating Graph for comparison 2018 respondent's birth cohort against marital status ####
cohort_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2018 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(text = element_text(size=6))
```

@@

#### Creating Graph for comparison 2016 respondent's birth cohort against marital status ####
cohort_marital_2016 <- ggplot(data=cleaned_gss_2016, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2016 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

@@

#### Creating Graph for comparison 2014 respondent's birth cohort against marital status ####
cohort_marital_2014 <- ggplot(data=cleaned_gss_2014, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Birth Cohort",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2014 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "2014-2021 Comparisons of Birth Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(cohort_marital_2021) + (cohort_marital_2018) + (cohort_marital_2016) + (cohort_marital_2014) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

@@

```{r}
#| include: false
#| message: false
#### Creating Graph for comparison 2021 respondent's sex against marital status ####
sex_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = sex, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Sex",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false
#| 
#### Creating Graph for comparison 2018 respondent's sex against marital status ####
sex_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = sex, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Sex",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "2014-2021 Comparisons of Respondent's Sex Against Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(sex_marital_2021) + (sex_marital_2018) + (sex_marital_2016) + (sex_marital_2014) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

## Marital Status and Self-Ascribed Belonging to Political Parties

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "2014-2021 Comparisons of Political Party Identity Against Respondent's Marital Status"
#### Using Patchwork Package to Combine Two Graphs ####
(partyid_marital_2021) + (partyid_marital_2018) + (partyid_marital_2016) + (partyid_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-4
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2021 Respondents"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-5
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2018 Respondents"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-6
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2016 Respondents"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-7
#| tbl-cap: "Marital Status vs. Self-Ascribed Belonging to Political Parties of 2014 Respondents"

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-4
#| fig-cap: "2014-2021 Comparisons of Birth Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine four Graphs ####
(polviews_marital_2021) + (polviews_marital_2018) + (polviews_marital_2016) + (polviews_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

@@

\appendix

>>>>

variable_table <-
  tibble(
    "Variable" = c(
      "partyid",
      "polviews",
      "marital",
      "sex",
      "cohort"
    ),
    "Variable Description" = c(
      "Self-described identification with a political party",
    "Self-described placement on the political spectrum",
    "Marital status of respondent",
    "Self-described sex of respondent",
    "Generational cohort that respondent belongs to "
    )
  )

@@

### Marital Status and Generational Cohort

Another important point of bias to acknowledge is the effect of generational cohort on marital status. As can be seen in @fig-1, the effects of age on generational cohorts largely impacts the marital status. For example, in 2014, there are significantly more respondent's belonging to the "Post-War" cohort and no respondent's belonging to the "Gen Z" cohort. This is influential as the 2014 data is missing marital rates from the "Gen Z" cohort as they were not old enough to be surveyed, thereby distorting marital rates in that year when compared to the follow three years. Additionally, due to the young age of "Gen Z", respondents many have not reached the age where they are looking for long-term relationships resulting in very few being married and even less being divorced and widowed, as shown in @fig-1. However, it is important to include them in this analysis as they are a growing and extremely relevant cohort in demographic studies. Additionally, it is noticeable that trends in marriage between generational cohorts are shifting as can be seen by the increase of "Never Married" respondents between cohorts each year. This is significant as it may act as a co-indicator of marital rates in the U.S.

```{r}
#| include: false
#| message: false
#### Creating Graph for comparison 2021 respondent's generational cohort against marital status ####
cohort_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Generational Cohort",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

```{r}
#| include: false
#| message: false
#### Creating Graph for comparison 2018 respondent's generational cohort against marital status ####
cohort_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "enerational Cohort",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2018 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(text = element_text(size=6))
```

@@

#### Creating Graph for comparison 2016 respondent's generational cohort against marital status ####
cohort_marital_2016 <- ggplot(data=cleaned_gss_2016, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Generational Cohort",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2016 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

@@

#### Creating Graph for comparison 2014 respondent's generational cohort against marital status ####
cohort_marital_2014 <- ggplot(data=cleaned_gss_2014, aes(x = cohort, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Generational Cohort",
    y = "# of Respondents",
    fill = "Marital Status",
    caption = "2014 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") +
  theme(legend.position = "none") +
  theme(text = element_text(size=6))
```

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "2014-2021 Comparisons of Generational Cohort Against Respondent Marital Status"
#### Using Patchwork Package to Combine four Graphs ####
(cohort_marital_2021) + (cohort_marital_2018) + (cohort_marital_2016) + (cohort_marital_2014) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

@@

#### Creating Graph for comparison 2021 respondent's sex against marital status ####
sex_marital_2021 <- ggplot(data=cleaned_gss_2021, aes(x = sex, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Sex",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(text = element_text(size=6))
```

#### Creating Graph for comparison 2018 respondent's sex against marital status ####
sex_marital_2018 <- ggplot(data=cleaned_gss_2018, aes(x = sex, fill = marital)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(width =.8) +
  theme_classic() + 
  labs(
    x = "Sex",
    y = "# of Respondents ",
    fill = "Marital Status",
    caption = "2021 GSS"
  ) +
  scale_fill_brewer(palette="Spectral") + 
  theme(legend.position = "none")+
  theme(text = element_text(size=6))
```

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "2014-2021 Comparisons of Respondent's Sex Against Marital Status"
#### Using Patchwork Package to Combine four Graphs ####
(sex_marital_2021) + (sex_marital_2018) + (sex_marital_2016) + (sex_marital_2014) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

## Marital Status and Self-Described Belonging to Political Parties

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "2014-2021 Comparisons Respondent's Marital Status vs. Identification with a Political Party"

#### Using Patchwork Package to Combine Two Graphs ####
(partyid_marital_2021) + (partyid_marital_2018) + (partyid_marital_2016) + (partyid_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-4
#| tbl-cap: "2021 Respondent's Marital Status vs. Self-Described Identification with a Political Parties"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-5
#| tbl-cap: "2018 Respondent's Marital Status vs. Self-Described Identification with a Political Parties"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-6
#| tbl-cap: "2016 Respondent's Marital Status vs. Self-Described Identification with a Political Parties"

@@

```{r}
#| echo: false
#| warning: false
#| label: tbl-7
#| tbl-cap: "2014 Respondent's Marital Status vs. Self-Described Identification with a Political Parties"

@@

```{r}
#| echo: false
#| warning: false
#| label: fig-4
#| fig-cap: "2014-2021 Comparisons of Self-Described Placement on the Political Spectrum Against Respondent Marital Status"
#### Using Patchwork Package to Combine four Graphs ####
(polviews_marital_2021) + (polviews_marital_2018) + (polviews_marital_2016) + (polviews_marital_2014)+ 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2)+
  plot_layout(guides = "collect")
```

@@
